---
title: "R Notebook"
output: html_notebook
---

0. Model definition
```{r}
# Using Latin Hypercube Sampling for Approximate bayesian computation
# Step 1: Create a nl object:

library(nlrx)
library(ggplot2)
library(dplyr)
library(docstring)

# Define paths
netlogopath <- "/home/jmtubay/Documents/NetLogo 6.2.2"
modelpath <- file.path(
  netlogopath, "app", "models", "Code Examples",
  "Extensions Examples", "gis", "swine_model_asf_v6_anon_WB_ONLY.nlogo"
)
modelname <- basename(modelpath)
outpath <- "out/"

# Check if paths exist
check_paths <- function(netlogopath, modelpath, outpath) {
  #' @title Check paths
  #' @description This function check the paths of the netlogo installation,
  #' the netlogo model file, and the output folder
  #' @param netlogopath Path of the NetLogo installation
  #' @param modelpath Path of the NetLogo model file
  #' @param outpath Path of the output directory for nlrx
  #' @return Prompts whether the paths are found or not.

  if (dir.exists(netlogopath)) {
    message(paste("NetLogo path found:", netlogopath))
  } else {
    stop(paste("NetLogo path not found:", netlogopath))
  }

  # Check if model file exists
  if (file.exists(modelpath)) {
    message(paste("Model file found:", modelpath))
  } else {
    stop(paste("Model file not found:", modelpath))
  }

  # Check if output directory exists, create if it doesn't
  if (dir.exists(outpath)) {
    message(paste("Output directory exists:", outpath))
  } else {
    warning(paste("Output directory does not exist. Creating:", outpath))
    dir.create(outpath, recursive = TRUE)
  }
}

check_paths(netlogopath, modelpath, outpath)

# Define Java options to suppress warnings
options(java.parameters = "-Djava.util.logging.ConsoleHandler.level=SEVERE")

# Initialize nlrx object
nl <- nl(
  nlversion = "6.2.2",
  nlpath = netlogopath,
  modelpath = modelpath,
  jvmmem = 1024
)

# Print nlrx object to verify
print(nl)

# Step 2: Attach an experiment
# We want to run a Latin Hypercube sampling, thus we need to define proper
# variable ranges. We also need to define our output metrics. These metrics are
# also used for the rejection sampling later on.

nl@experiment <- experiment(expname="asf-model-nl",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="false",
                            idsetup="setup-all-in",
                            idgo="go",
                            runtime=53,
                            metrics=c("avg-distance-traveled-by-infection"),
                            variables = list(#"num-patches-to-infect" = list(min=0, max=20, step=1, qfun="qunif"),
                                             #"alpha-farm-farm" = list(min=0, max=20, step=1, qfun="qunif"),
                                             #"alpha-farm-patch" = list(min=0, max=20, step=1, qfun="qunif"),
                                             #"alpha-patch-farm" = list(min=0, max=20, step=1, qfun="qunif"),
                                             "alpha-patch-patch" = list(min=0, max=3, step=0.1, qfun="qunif"),
                                             #"beta-farm-farm" = list(min=0, max=20, step=1, qfun="qunif"),
                                             #"beta-farm-patch" = list(min=0, max=20, step=1, qfun="qunif"),
                                             #"beta-patch-farm" = list(min=0, max=20, step=1, qfun="qunif"),
                                             "beta-patch-patch" = list(min=0, max=3, step=0.1, qfun="qunif")),
                            constants = list("num-patches-to-infect" = 1,
                                             "termination-t" = 53,
                                             "cull?" = "false",
                                             "cull-infected-farm?" = "false",
                                             "stop-at-t?" = "true",
                                             "infect-random-wildboar-area-per-time-step?" = "false",
                                             "cull-n-random-wildboar-area?" = "false",
                                             "restrict-farm-connection?" = "false",
                                             "stop-when-I-plateaus?" = "false"
                            ),
)

# Step 3: Attach a simulation design
# While the experiment defines the variables and specifications of the model,
# the simulation design creates a parameter input table based on these model
# specifications and the chosen simulation design method. nlrx provides a bunch
# of different simulation designs, such as full-factorial, latin-hypercube,
# sobol, morris and eFast (see “Simdesign Examples” vignette for more
# information on simdesigns). All simdesign helper functions need a properly
# defined nl object with a valid experiment design. Each simdesign helper also
# allows to define a number of random seeds that are randomly generated and can
# be used to execute repeated simulations of the same parameter matrix with
# different random-seeds (see “Advanced configuration” vignette for more
# information on random-seed and repetition management). A simulation design is
# attached to a nl object by using one of the simdesign helper functions:

samples <- 500
nseeds <- 10
precision <- 3
nl@simdesign <- simdesign_lhs(
  nl = nl,
  samples = samples,
  nseeds = nseeds,
  precision = precision
)
```


1. Load the Results RDS File:
First, read the saved .rds file into your R session to retrieve the nl object and its associated simulation outputs.
```{r}
library(nlrx)
library(ggplot2)
library(dplyr)
library(docstring)

# Define the path to the saved RDS file
out <- "/home/jmtubay/Documents/ASF_Model/out/"
rds_path <- file.path(out, "ABClhs_swine_model_asf_v6_anon_WB_ONLY.nlogo_samples10000_seeds1_2025-02-23_05-43-59.rds")
# Extract the file name
rds_file_name <- basename(rds_path)

# Check if the RDS file exists
if (file.exists(rds_path)) {
  # Reload the saved nl object
  nl <- readRDS(rds_path)
  
  # Extract the results
  results <- getsim(nl, "simoutput")
  
  # Extract input
  input <- getsim(nl, "siminput")
  
} else {
  stop("The RDS file does not exist. You may need to rerun the simulations.")
}

print(nl)
print(input)
print(results)
```

2. Prepare Data for ABC:
Use the loaded nl object to extract input and output data, which will be used as parameters and summary statistics for the ABC methods.

```{r}
library(dplyr)

# Attach output results to nl objects
setsim(nl, "simoutput") <- results

# Extract input parameters
input <- getsim(nl, "siminput") %>%
  dplyr::select(names(getexp(nl, "variables")))

print(input)

# Extract output metrics
output <- getsim(nl, "simoutput") %>%
  dplyr::select(getexp(nl, "metrics"))

print(output)

nseeds = 1

# Check if 'simulation_id' column exists, if not, create it
if (!"simulation_id" %in% colnames(output)) {
  # Compute the correct number of simulation ids based on the total number of rows in 'output'
  total_rows <- nrow(output)
  
  # Create a simulation_id that corresponds to the total number of rows
  output <- output %>%
    mutate(simulation_id = rep(1:ceiling(total_rows / nseeds), each = nseeds, length.out = total_rows))
}

ModeNonZero <- function(x, na.rm = FALSE) {
  if (na.rm) {
    x <- na.omit(x)
  }
  x <- x[x != 0]  # Exclude zero elements
  if (length(x) == 0) {
    return(0)  # Return zero if there are no nonzero elements
  }
  ux <- unique(x)
  mode_value <- ux[which.max(tabulate(match(x, ux)))]
  if (is.na(mode_value)) {
    return(0)  # Replace NA with zero if mode is NA
  }
  return(mode_value)
}

MeanNonZero <- function(x, na.rm = FALSE) {
  if (na.rm) {
    x <- na.omit(x)
  }
  x <- x[x != 0]  # Exclude zero elements
  if (length(x) == 0) {
    return(0)  # Return zero if there are no nonzero elements
  }
  return(mean(x))  # Calculate mean of nonzero elements
}

# Summarize the output
output_summary <- output %>%
  group_by(simulation_id) %>%
  summarize(avg_metric = mean(`avg-distance-traveled-by-infection`, na.rm = TRUE)) %>%
  ungroup()

# Check the result
print(output_summary)

# Define target summary statistics
target <- c("avg_distance" = 30.66)
# https://doi-org.utrechtuniversity.idm.oclc.org/10.1038/s41598-023-42300-0 (0.042 km/day)
```

3. Perform ABC Rejection and Loclinear Methods:
Use the abc package to perform both the rejection and loclinear ABC methods with the prepared data.

```{r}
avg_metric <- output_summary %>% select(avg_metric)

# Perform ABC using rejection method
results.abc.reject <- abc::abc(
  target = target,
  param = input,
  sumstat = avg_metric,
  tol = 0.01,
  method = "rejection"
)


# Perform ABC using loclinear method
results.abc.loclin <- abc::abc(
    target = target,
    param = input,
    sumstat = output_summary$avg_metric,
    tol = 0.01,
    method = "loclinear"
  )

# Perform ABC using neural networks
results.abc.neuralnet <- abc::abc(
  target = target,
  param = input,
  sumstat = output_summary$avg_metric,
  tol = 0.01,
  method = "neuralnet"
)

print(results.abc.reject)
print(results.abc.loclin)
print(results.abc.neuralnet)
# If you need to plot, you can now check the content of results.abc.loclin
```

```{r}
# Assuming 'input' is a data frame or matrix used in the abc::abc() call
total_simulations <- nrow(input)  # or length(output_summary$avg_metric)

# Print the total number of simulations
print(total_simulations)
```

Theme
```{r}
library(remotes)

remotes::install_git("https://gitlab.envt.fr/epidesa/biosecureR.git", host = "gitlab.envt.fr")
```


4. Compare and Visualize Results:
Format the results for visualization and use ggplot2 to create plots comparing the parameter distributions from the ABC methods and the initial distribution from Latin Hypercube Sampling (LHS).

```{r}
# Create an empty list to store available results
results_list <- list()

# Check each method and add it to the results list if it's not NULL
if (!is.null(results.abc.reject$unadj.values)) {
  results_list$rejection <- tibble::as_tibble(results.abc.reject$unadj.values) %>%
    tidyr::gather(parameter, value) %>%
    dplyr::mutate(method = "rejection")
}

if (!is.null(results.abc.loclin$adj.values)) {
  results_list$loclinear <- tibble::as_tibble(results.abc.loclin$adj.values) %>%
    tidyr::gather(parameter, value) %>%
    dplyr::mutate(method = "loclinear")
}

if (!is.null(results.abc.neuralnet$adj.values)) {
  results_list$neuralnet <- tibble::as_tibble(results.abc.neuralnet$adj.values) %>%
    tidyr::gather(parameter, value) %>%
    dplyr::mutate(method = "neuralnet")
}

# Always include the initial parameter distribution
#results_list$lhs <- input %>%
#  tidyr::gather(parameter, value) %>%
#  dplyr::mutate(method = "lhs (init dist)")

# Combine all available results into a single tibble
results.abc.all <- dplyr::bind_rows(results_list)

# Calculate the mean for each parameter-method combination
means <- results.abc.all %>%
  dplyr::group_by(parameter, method) %>%
  dplyr::summarise(mean_value = mean(value, na.rm = TRUE), .groups = "drop")

# Calculate the mode (peak) using density estimation
find_peak <- function(values) {
  d <- density(values, na.rm = TRUE)
  peak_x <- d$x[which.max(d$y)]
  return(peak_x)
}

peaks <- results.abc.all %>%
  dplyr::group_by(parameter, method) %>%
  dplyr::summarise(peak_value = find_peak(value), .groups = "drop")

# Print means and peaks
print("Mean and peak values for each parameter-method combination:")
results_summary <- dplyr::left_join(means, peaks, by = c("parameter", "method"))
print(results_summary)

# Plotting
plot_density <- ggplot2::ggplot(results.abc.all) +
  ggplot2::facet_wrap(~parameter, scales = "free") +
  ggplot2::geom_density(ggplot2::aes(x = value, fill = method), alpha = 0.1, size = 1) +
  ggplot2::geom_vline(
    data = means,
    ggplot2::aes(xintercept = mean_value, color = method),
    linetype = "dashed", size = 1.2
  ) +
  ggplot2::geom_vline(
    data = peaks,
    ggplot2::aes(xintercept = peak_value, color = method),
    linetype = "solid", size = 1.2
  ) +
  ggplot2::xlim(-1, 4) + # Adjust x-axis range as needed
  ggplot2::labs(title = "", x = "Value", y = "Density", fill = "Legend Title", color = "Legend Title") +
  ggplot2::guides(fill = ggplot2::guide_legend(title = "ABC Methods"), color = ggplot2::guide_legend(title = "ABC Methods")) +
  biosecureR::scale_fill_biosecure() + # Apply Biosecure fill scale
  biosecureR::scale_color_biosecure() + # Apply Biosecure color scale
  biosecureR::theme_biosecure() + # Apply Biosecure theme
  ggplot2::theme(
    #legend.position = "bottom",  # Place legend at bottom for better visibility
    panel.border = ggplot2::element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank()
  )


# Save the plot
ggsave("param_est.png", width = 10, height = 6, dpi = 300)

# Show the plot
print(plot_density)
```
